---
title: "Xaringan Template"
subtitle: "An example"
author: "Patricia Menéndez"
institute: "Department of Econometrics and Business Statistics"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: css/style.css
    yolo: FALSE
---


```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r , echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(dplyr)
library(magrittr)
library(ggplot2)
```



---
class: center, middle
# Hello!

---
class: left, top
# This is an example of using boxes around txt
.content-box-soft[ 
- Understand the aims and structure of the unit
]

.content-box-duke-green[ 
* Explain the need for reproducible and collaborative practices
]
.content-box-soft[ 
* Introduce the main tools that we are going to use during the lectures/tutorials
]

---
class: left, top

# Adding color to the text

- .green[ This is an example in green]
- .red[ This is an example in green]

- .bold[This is an example in bold]

- you can also use markdown bold syntax **bold**

---

class: left, middle
# Using also LaTex

.content-box-soft[ 
- .green[.bold[Patricia]] $\leftarrow$ hello hello ]

- $\sum_{i = 1}^{\infty} x_i$

**Bolding using markdown**

---
class: left, top

# research question 

-1. How is the trend of **malaria** change across the years in the WORLD ? 

-2. What is the 5 countries with highest death rates for malaria

-3. What is the preventive measures for malaria? 


---
class: left, top

# Malaria world trend 

```{r, echo = FALSE, preserve = TRUE}

malaria_data <- read.csv("Data/annual-number-of-deaths-by-cause (1).csv")

```

.pull-left1[
```{r, echo = FALSE, preserve = TRUE, fig.cap ="Malaria world trend"}

malaria_data <- malaria_data %>%
  select(Entity, Year, Deaths...Malaria...Sex..Both...Age..All.Ages..Number.) %>%
  rename(Malaria = Deaths...Malaria...Sex..Both...Age..All.Ages..Number.) %>%
  rename(Country = Entity) %>%
  filter(Year >= 2000) 

malaria_world<- malaria_data %>%
  filter(Country == "World") 

ggplot(malaria_world, aes( x = Year, y = Malaria)) +
  geom_line()+
  scale_y_continuous(labels = scales::comma)




```
]


- The world trend of Malaria death rate was **decreasing** from 2000 to 2019


---
class: left, top

# Top 5 Countries with **highest** Malaria death rate 

```{r, echo = FALSE, preserve = TRUE}

malaria_mean <- malaria_data %>%
  group_by(Country) %>%
  summarize(Mean = mean(Malaria)) %>%
  mutate(Mean = ceiling(Mean)) %>%
  arrange(-Mean) %>%
  filter(Country != "World") %>%
  filter(Country != "Sub-Saharan Africa (WB)") %>%
  filter(Country != "African Region (WHO)") %>%
  filter(Country != "World Bank Lower Middle Income") %>%
  filter(Country != "World Bank Low Income") %>%
  filter(Country != "Democratic Republic of Congo") %>%
  filter(Country != "South-East Asia Region (WHO)") %>%
  filter(Country != "South Asia (WB)") %>%
  filter(Country != "G20") 

head(malaria_mean, 5)

#knitr::kable(caption = "Top 5 Countries with highest Malaria death rate")


```



.pull-left1[
```{r, echo = FALSE, preserve = TRUE, fig.cap ="Malaria world trend"}

malaria_top5 <- malaria_data %>%
  filter(Country %in% c("Nigeria", "India", "Uganda", "Burkina Faso", "Cote d'Ivoire")) 

ggplot(malaria_top5, aes( x = Year, y = Malaria, color = Country)) +
  geom_line()+
  scale_y_continuous(labels = scales::comma)




```
]

-The trend of malaria death rate from 2000 to 2019 was **decreasing** for the top 5 countries


---
class: left, top

# Reasons for high death rate of malaria in Nigeria, India, Uganda, Burkina Faso, Cote d'Ivoire		 






---
class: left, top

# Preventive measures for Malaria








---
class:  left, top

# Adding figure

.pull-left1[
```{r  penguins , eval = FALSE, warning = FALSE, message = FALSE}
ggplot(data = penguins, 
       aes(x = flipper_length_mm,
           y = body_mass_g)) +
  geom_point(aes(color = species,
                 shape = species),
             size = 2) +
  scale_color_manual(
    values = c("darkorange",
               "darkorchid",
               "cyan4"))

```
]


.pull-right1[
```{r  penguins2 , eval = TRUE, echo = FALSE, warning = FALSE, message = FALSE}
ggplot(data = penguins,
       aes(x = flipper_length_mm,
           y = body_mass_g)) +
  geom_point(aes(color = species,
                 shape = species),
             size = 2) +
  scale_color_manual(
    values = c("darkorange",
               "darkorchid",
               "cyan4"))

```
]



---
class:  left, top
# Adding figure --> another way

.right-column[
```{r  penguins3 , eval = FALSE, warning = FALSE, message = FALSE}
ggplot(data = penguins, 
       aes(x = flipper_length_mm,
           y = body_mass_g)) +
  geom_point(aes(color = species,
                 shape = species),
             size = 2) +
  scale_color_manual(
    values = c("darkorange",
               "darkorchid",
               "cyan4"))

```
]


.left-column[
```{r  penguins4 , eval = TRUE, echo = FALSE, warning = FALSE, message = FALSE}
ggplot(data = penguins,
       aes(x = flipper_length_mm,
           y = body_mass_g)) +
  geom_point(aes(color = species,
                 shape = species),
             size = 2) +
  scale_color_manual(
    values = c("darkorange",
               "darkorchid",
               "cyan4"))

```
]




---

```{r endslide, child="EndSlide/endslide.Rmd"}
```  

